---
title: "Introduction to Statistical Learning Chapter 5 Problem 9"
author: "Robert Smith"
date: "7/30/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Exercises

##Applied

(9) We will now consider the Boston housing data set, from the MASS library.

(a) Based on this data set, provide an estimate for the population mean of medv.
```{r,echo=T}
library(MASS)
data(Boston)
uhat<-mean(Boston$medv)
print(uhat)
```

(b) Provide an estimate of the standard error of uhat
```{r,echo=T}
SE<-sd(Boston$medv)/sqrt(nrow(Boston))
print(SE)
```

(c) Now estimate the standard error of uhat using the boostrap. How does this compare to your answer from (b)?
```{r,echo=T}
library(boot)
set.seed(12345)
boot.fn<-function(data,index) mean(Boston$medv[index])
bootstp<-boot(Boston,boot.fn,1000)
print(bootstp)
```
The bootstrap standard error is slightly lower than the estimated SE.

(d) Based on your bootstrap sample estimate from (c), provide 95% confidence intervals using the formula: [uhat - 2*SE(uhat),uhat + 2*SE(uhat)]
```{r,echo=T}
lower<-bootstp$t0 - 2*0.3997714
upper<-bootstp$t0 + 2*0.3997714
bounds<-c(lower,upper)
print(bounds)
```

Compare it to the results obtained using t.test(Boston$medv)
```{r,echo=T}
t.test(Boston$medv)
```
Confidence intervals are within first decimal accuracy.

(e) Based on this data set, provide an estimate uhat_med, for the median value of the medv in the population.
```{r,echo=T}
uhat_med<-median(Boston$medv)
print(uhat_med)
```

(f) We now would like to estimate the standard error of uhat_med. Unfortunately there is no simple formula for computing the standard error of the median. Instead, estimate the standard error of the median using the bootstrap. Comment on your findings.

```{r,echo=T}
set.seed(12345)
boot.fn<-function(data,index) median(Boston$medv[index])
bootstp<-boot(Boston,boot.fn,1000)
print(bootstp)
```
The median and bootstrap value are the same. 

(g) Based on this data set, provide an estimate for the tenth percentile of the medv in Boston suburbs. Call this quantity uhat_0.1. 
```{r,echo=T}
uhat_0.1<-quantile(Boston$medv,0.1)
print(uhat_0.1)
```

(f) Use the bootstrap to estimate the standard error of uhat_0.1. Comment on your findings.
```{r,echo=T}
set.seed(12345)
boot.fn<-function(data,index) quantile(Boston$medv[index],0.1)
bootstp<-boot(Boston,boot.fn,1000)
print(bootstp)
```
The quantile and bootstrap estimate are the same.